\textbf{Access Control for Content Distribution Network Assets}

### 2016-05-11 tex/paper.tex
Changed according to reviewer comments

%}}}
\section{Introduction}
Developing large projects containing static content usually involves using a Content Distribution
Network to be able to scale to a larger user base. The commercial Content Distribution Networks are
usually fairly easy to use. The content that is to be used in a project is uploaded and then
distributed over the globe when a client is trying to access the content. For secret content this
can be a problem and an inconvenience, and that is what this thesis is about.

@@

As large projects most often are divided into different dividable parts, a concept called containers
is introduced in this work. Containers can contain more containers and content, and when
modifying the security settings for a container it is applied recursively to all its content and
sub-containers.

This work examines a way of enforcing virtual access control on content and groups of content, in
the form of containers and snapshots to be able to handle the secrecy requirements easily. A system
was developed using the Copy-on-Write principle to improve scalability and make it a lighter
operation to take snapshots of parts of the system.

The research question that this report tries to answer is how and whether it is practically 
feasible to use Copy-on-Write for a high-level system like the one that is implemented and how 
it can effect the system's scalability.

\newpage
\section{Background}
This section gives the reader some background of the company where the thesis was written and
implementation was made. It also gives some brief explanations of the core concepts which are 
exerted throughout this work.

@@

\subsubsection{Snapshots}
Several filesystems~\cite{BTRFS}\cite{ZFS} use snapshot techniques to make it possible for a user
or system to roll back the filesystems state, to a state that the filesystem had at the time of
snapshot creation. This can be useful for example when having a stable system and then doing a
system update, if the system does not respond well to the update the filesystem can be rolled
back to its previous stable state, provided that a snapshot was created before the update procedure.

This work has taken a lot inspiration from the way BTRFS~\cite{BTRFS} handles snapshots, which is
further explained in Section~\ref{sec:btrfs}.

\newpage
\subsection{Related Terminology}
In this section concepts and abbreviations, that are recurring throughout the paper, which the
reader needs to be familiar with are explained.

@@

\subsubsection{Terms}
\begin{description}
\item[Eventual Consistency] \hfill \\
Eventual consistency is an informal guarantee that defines how the consistency of stored data can
behave. In a system defined as using eventual consistency, access to stored data is not guaranteed
to return the result of the latest modification. At some point, if modifications to the stored data
stop, access to a data item will return the resulting data of the last modification of that
item.~\cite{EVENTUAL}. It is often used in the context of database systems.

@@

\begin{center}
    \begin{tabular}{ | l | l | l | p{5cm} |}
        \hline
        \textbf{Set} & \textbf{Relation} & \textbf{Type} \\ \hline
        \textbf{C} & $\textbf{C} \supseteq C \ni c$ & $\textbf{C}=P_{fin}(P_{fin}(C \cup M))$\\ \hline
        \textbf{M} & $\textbf{M} \supseteq M \ni m$ & $\textbf{M}=P_{fin}(M)$\\ \hline
        \textbf{F} & $\textbf{F} \supseteq F \ni f$ & $\textbf{F}=P_{fin}(F)$\\ \hline
    \end{tabular}
\end{center}

@@

\begin{center}
    \begin{tabular}{ | l | l | p{5cm} |}
        \hline
        \textbf{Element} & \textbf{Type} \\ \hline
        c & $C \cup M$ \\ \hline
        m & $\{id: GUID, file: f\}$ \\ \hline
        f & File \\ \hline
    \end{tabular}
\end{center}

@@

\item[Access Matrix] \hfill \\
    The access matrix contains what containers that all the users in the system have access to. If
    A[u,c] is true in the access matrix, user \textit{u} has access to container \textit{c}.

@@

\subsection{Access rights}
Access is only handled on containers and if a user has access to a container it also has access to
all of that containers children.
The \textit{A[u,c]} notation is symbolising an access token. If \textit{A[u,c]} is true in the 
access matrix \textit{A}, user \textit{u} has full access to the container object \textit{c} and its
children. If it is false or if the entry does not exist in the access matrix, the user \textit{u}
does not have any access rights for \textit{c}.

\begin{equation}
    \begin{split}
        A[u,c] \Rightarrow & u \text{ can read } c \\
        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow & u \text{ can write to } c \\
        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow & u \text{ can delete } c \\
        A[u,c] \land f \in c \centernot\Rightarrow & u \text{ can delete } f 
    \end{split}
\end{equation}

\subsection{Data integrity}
A computer system or subsystem is defined as possessing the property of integrity if it behaves consistently
according to a defined standard. This implies that a subsystem possessing the property of integrity
does not guarantee an absolute behaviour of the system, but rather that it performs according to
what its creator intended ~\cite{BIBA}.
\\
The invariant for this model is the file. Which means that it should not be possible to
change a file though out the cycle of operation without creating a new file object with a new
identifier. This is to ensure that files can not lose their integrity by concurrent
operations from different nodes or get mixed up when referred to from a content object.

@@

\subsection{Integrity Threats}
According to Biba et. al~\cite{BIBA} one can consider two threat sources, namely subsystem external
and subsystem internal. The external sources could be another system calling the subsystem with
faulty data or trying to make inaccurate calls to program functions. Another external source could
also be somebody trying to tamper with the exposed functions of the program. Threats that are
internal could be a malicious part of the subsystem or simply an incorrect part of the subsystem,
which does not behave according to its specification.

@@

\subsection{Consequences of modification} \label{sec:conseq}

In this section the consequences, side effects and direct effects, of different operations are
stated in an informal mathematical manner, similar to how inference rules are composed. These rules
are what should be expected to be followed by the implementation in
Section~\ref{sec:implementation}.

@@

\vspace{2em}
\textbf{File Read}
\vspace{1em}
\begin{equation} \label{eq:fileread}
    \inferrule
    {(\textbf{*}) \\ C \ni c \ni m \\ A[u,c]} 
    {(C, F) \xrightarrow{u, read(m,f)} (C, F)}
\end{equation}
\vspace{1em}

A user \textit{u} with access rights to a parent container \textit{c} of content \textit{m}
can read a file \textit{f} in content \textit{m}.

The predicates in (\ref{eq:fileread}) are used in several places and are therefore named \textit{*}.

\vspace{2em}
\textbf{File Creation}
\vspace{1em}
\begin{equation} \label{eq:filecreate}
    \inferrule
    {C \ni c \\ A[u,c] \\ m \centernot \in c \\ \neg readOnly(c)} 
    {(C \cup \{c\}, F) \xrightarrow{u, create(m,c)} (C \cup \{c \cup \{m\}\}, F \cup \{m.file\})}
\end{equation}
\vspace{1em}

A user \textit{u} with access rights to a non-readonly container \textit{c} can create a content
\textit{m}, containing a file \textit{f}, in container \textit{c}.

@@

If a user \textit{u} with access rights to the parent (source and destination) container \textit{c}
of content \textit{m} copies that content, the new content \textit{m'} is not equal to the old
(because of its new ID), however they both contain the same file.

@@

\vspace{2em}
\textbf{Snapshot}
\vspace{1em}
\begin{equation} \label{eq:snapshot}
    \begin{split}
        [C', M'] := snapshot(tree) & \Implies \\
        \forall c \in tree, copy(c) \in & C' \\
        \forall m \in tree, copy(m) \in & M' \\
        \forall f \in tree, f \in & F
    \end{split}
\end{equation}

To explain the non-trivial snapshot operation it is first defined in \ref{eq:snapshot}. The snapshot
operation is performed upon a tree of objects and results in two new sets (C' and M'), one with
containers and one with content, the referred files remain the same as before the snapshot.

@@

\subsection{Interaction of Multiple Accesses} \label{sec:multiple_access}
As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} in
Section~\ref{sec:conseq} there can be race conditions where the last write wins. This could be
solved by locks or transactions, but as updates are not based on each other, the trade-off for this
inconsistency to scalability and speed is intentional. However, as can also be seen in
consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
the set of files, even ough its modification process is the last one to write to the content.
This is necessary as a file can be referenced from several content and the removal process can not
atomically ensure that the file is not referred to by another content. This results in a lot of
files in the system which are not referenced from anywhere. This can be solved by either making each
file keep track of how many times it is referred to and also make it possible to check this number
and delete the file in an atomic fashion. An easier approach would be to have a garbage collector
which locks parts of the system momentarily meanwhile it removes files which are not referenced.

\par
There is also the chance to have content and containers that are not referenced from anywhere and
result in being garbage in the system. This could happen as a result of consequence~\ref{eq:delete}
in combination with any of the other consequences listed in Section~\ref{sec:conseq}. The basic idea
is that an entity is deleted meanwhile another is created or modified to have that entity as its
parent, thus creating a separate tree that can not be reached from the root node of the project.
These type of inconveniences should also be cleaned up by the garbage collector.

\subsection{JPF}
JPF was used to test the idea and state transitioning of the model. It is a very simplified
version of the real system that still contains all the important Copy-on-Write concepts and the 
assumptions that have been made in the model. This simplified version was then used to automatically
test for unwanted quirks. It is not a proof that the model works, but it is very exhaustive in its
testing.

@@

\subsubsection{Garbage Collection}
For the system to reach a clean state and the decided eventual consistency~\cite{KLINGSBO}, garbage
collection will be needed. As of the conclusion in Section~\ref{sec:multiple_access} from Section
~\ref{sec:conseq} there will be objects in the system that are unreferenced and can not be reached
and should therefore be cleaned up for them not to cause negative performance effects on the system.

@@

\newpage
\section{Summary}
This section covers the conclusions that could be drawn from this work and how the implementation 
could, and will be, extended in the future.

@@

\par If this application would be distributed with several persistent storage nodes like MongoDB,
the application would not always be in a global consistent state as there would not be any global
locks.  This could theoretically cause some inconvenience for the user but in all real world
observations no users have noticed it. The theoretical inconvenience for the user was a trade-off
made so that the application could scale on the width almost endlessly, especially if the issue with
building the project trees is solved (as mentioned in Section~\ref{sec:load_testing}).

@@

\par The conclusion that can be drawn from implementing Copy-on-Write as high up in the stack as it
was in Perius is that the code would have to be largely rewritten for every application that wants
to use a similar approach. The solution to this could be to use Copy-on-Write in the persistent
storage instead of in the actual application, which also efficiently could handle conflict 
resolution and every improvement made to it could be shared for every system that uses the same 
type of system architecture and underlying persistent storage. On the other hand, by having all the
Copy-on-Write specific code in the back-end instead of the persistent storage the choices of
storage becomes wider as the database does not have to implement the named things. This could be
positive when deploying Perius in environments with high requirements for the number of users
that the system should be able to handle. It would also be well suited if the level of consistency
needs to be fully tweakable to make the trade off for how many users the application would work for
and consistency of data.

\subsection{Future work}
\subsubsection{Access Control}
Full access control was not implemented according to the model described in ~\ref{sec:model}, it
was only implemented to check whether a user should have access to the system as a whole or not, the
implementation did not set or check any specific access rights to certain contents or containers.

\subsubsection{Front-end Refactorization}
The application could be made substantially more efficient by rewriting the front-end to update
itself according to the REST response after a modification of the project tree, instead of 
re-fetching the full project tree every time a change is made. The front-end should also be fixed to
properly follow the Reflux streamlining.

@@

%{{{

%What to write about
%Battle Binary
%Security features of CDNn's
%What is needed by the logic, like what happens after two consecutive deletes
%
%What is needed
%* Security layers
%* snapshots
%* Virtual file structure
%* Versioning of content
%* Multi project support
%* Auth and audit logs
%* Users

%Stuff to write about:
%Modular design, every piece should be interchangeable
%LDAP - why it was used as standard AUTH

%Using copy-on-write as the sole update strategy has pros and cons. The upside is
%that it is simple to guarantee operation atomicity, and data-structure integrity. The
%downside is that performance relies on the ability to maintain large extents of free
%contiguous disk areas. In addition, random updates to a f i le tend to fragment it,
%destroying sequentiality. A good defragmentation algorithm is required; this is
%described in Section 5

%}}}

>>>>

%}}}
\section{Introduction}
The development of large projects using static content usually involves the use of a Content
Distribution Network to be able to scale to a larger user base. Commercial Content Distribution
Networks are usually fairly easy to use. The content that is to be used in a project is uploaded and
then distributed over the globe when a client is trying to access the content. For secret content
this can be a problem and an inconvenience due to the lack of tools available to handle groups of
content and their security settings, and that is what this thesis is about.

@@

As large projects most often are divided into parts, a concept called containers is introduced in
this work. Containers can contain more containers and content, and when modifying the security
settings for a container it is applied recursively to all its content and sub-containers.

This work examines a way of enforcing access control on content and groups of content, in the form
of containers and snapshots. By being able to control the secrecy settings on a container instead of
separately modifying the settings for each individual content it becomes easier to release content
to be accessible to the public or different groups during the progression of the project. To manage
different versions or releases of the project, the snapshot functionality can be used to handle
access control by taking a snapshot with the correct security settings at each release.

A system was developed using the Copy-on-Write principle to improve scalability and make it a
lighter operation to take snapshots of parts of the system.

The research question that this report tries to answer is how and whether it is practically feasible
to use Copy-on-Write for a high-level storage system and how it can affect the scalability of the
system.

\newpage
\section{Background}
This section gives the reader some background of the company where the thesis was written and the
implementation was made. It also gives some brief explanations of the core concepts which are used
throughout this work.

@@

\subsubsection{Snapshots}
Several filesystems~\cite{BTRFS}\cite{ZFS} use snapshot techniques to make it possible for a user
or system to roll back the system's state, to a state that the filesystem had at the time of
snapshot creation. This can be useful for example when having a stable system and then doing a
system update, if the system does not respond well to the update the filesystem can be rolled
back to its previous stable state, provided that a snapshot was created before the update procedure.

This work has taken a lot inspiration from the way the BTRFS filesystem~\cite{BTRFS} handles
snapshots, which is further explained in Section~\ref{sec:btrfs}.

\newpage
\subsection{Related Terminology}
In this section concepts and abbreviations that are recurring throughout the paper and which the
reader needs to be familiar with are explained.

@@

\subsubsection{Terms}
\begin{description}
\item[Eventual Consistency] \hfill \\
Eventual consistency~\cite{EVENTUAL} is an informal guarantee that defines how the consistency of
stored data can behave. In a system defined as using eventual consistency, access to stored data is
not guaranteed to return the result of the latest modification. At some point, if modifications to
the stored data stop, access to a data item will return the resulting data of the last modification
of that item. It is often used in the context of database systems.

@@

\begin{center}
    \begin{tabular}{ | l | l | l | p{5cm} |}
        \hline
        \textbf{Set} & \textbf{Relation} & \textbf{Type} \\ \hline
        \textbf{C} & $\textbf{C} \supseteq C \ni c$ & $\textbf{C}=P_\mathit{fin}(P_\mathit{fin}(C \cup M))$\\ \hline
        \textbf{M} & $\textbf{M} \supseteq M \ni m$ & $\textbf{M}=P_\mathit{fin}(M)$\\ \hline
        \textbf{F} & $\textbf{F} \supseteq F \ni f$ & $\textbf{F}=P_\mathit{fin}(F)$\\ \hline
    \end{tabular}
\end{center}

@@

\begin{center}
    \begin{tabular}{ | l | l | p{5cm} |}
        \hline
        \textbf{Element} & \textbf{Type} \\ \hline
        c & $C \cup M$ \\ \hline
        m & $\{\mathit{id}: \mathit{GUID}, \mathit{file}: f\}$ \\ \hline
        f & File \\ \hline
    \end{tabular}
\end{center}

@@

\item[Access Matrix] \hfill \\
    The access matrix contains what containers that all the users in the system have access to. If
    $A[u,c]$ is true in the access matrix, user \textit{u} has access to container \textit{c}.

@@

\subsection{Access rights}
Access is only handled on containers and if a user has access to a container it also has access to
all of that containers children.
The $A[u,c]$ notation is symbolising an access token. If $A[u,c]$ is true in the 
access matrix \textit{A}, user \textit{u} has full access to the container object \textit{c} and its
children. If it is false or if the entry does not exist in the access matrix, the user \textit{u}
does not have any access rights for \textit{c}.

\begin{equation}
    \begin{split}
        A[u,c] \Rightarrow{}& u \text{ can read } c \\
        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow{}& u \text{ can write to } c \\
        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow{}& u \text{ can delete } c \\
        A[u,c] \land f \in c \centernot\Rightarrow{}& u \text{ can delete } f 
    \end{split}
\end{equation}

\subsection{Data integrity}
A computer system or subsystem is defined as possessing the property of integrity if it behaves consistently
according to a defined standard. This implies that a subsystem possessing the property of integrity
does not guarantee an absolute behaviour of the system, but rather that it performs according to
what its creator intended ~\cite{BIBA}.
\\
The invariant for this model is the file. Which means that it should not be possible to
change a file through out the cycle of operation without creating a new file object with a new
identifier. This is to ensure that files can not lose their integrity by concurrent
operations from different nodes or get mixed up when referred to from a content object.

@@

\subsection{Integrity Threats}
According to Biba et al.~\cite{BIBA} one can consider two threat sources, namely subsystem external
and subsystem internal. The external sources could be another system calling the subsystem with
faulty data or trying to make inaccurate calls to program functions. Another external source could
also be somebody trying to tamper with the exposed functions of the program. Threats that are
internal could be a malicious part of the subsystem or simply an incorrect part of the subsystem
that does not behave according to its specification.

@@

\subsection{Consequences of Operations} \label{sec:conseq}

In this section the preconditions and effects of different operations are
stated in an informal mathematical manner, similar to how inference rules are composed. These rules
should be followed by the implementation in Section~\ref{sec:implementation}.

@@

\vspace{1em}
The following predicates are used in several places and are therefore named \textit{*} to avoid
unnecessary repetition.
\begin{equation} \label{eq:fileread}
    \inferrule
    {(\textbf{*}) \\ C \ni c \ni m \\ $A[u,c]$} 
    {}
\end{equation}
\vspace{1em}

\vspace{2em}
\textbf{File Read}
\vspace{1em}
\begin{equation} \label{eq:fileread}
    \inferrule
    {(\textbf{*})} 
    {(C, F) \xrightarrow{u, read(m,f)} (C, F)}
\end{equation}
\vspace{1em}

A user \textit{u} with access rights to a parent container \textit{c} of content \textit{m}
can read a file~\(f\) in content \textit{m}.

\vspace{2em}
\textbf{File Creation}
\vspace{1em}
\begin{equation} \label{eq:filecreate}
    \inferrule
    {C \ni c \\ $A[u,c]$ \\ m \centernot \in c \\ \neg readOnly(c)} 
    {(C \cup \{c\}, F) \xrightarrow{u, create(m,c)} (C \cup \{c \cup \{m\}\}, F \cup \{m.file\})}
\end{equation}
\vspace{1em}

A user \textit{u} with access rights to a non-readonly container \textit{c} can create a
content~\(m\), containing a file \textit{f}, in container \textit{c}.

@@

If a user \textit{u} with access rights to the parent (source and destination) container~\(c\)
of content \textit{m} copies that content, the new content \textit{m'} is not equal to the old
(because of its fresh ID), however they both contain the same file.

@@

\vspace{2em}
\textbf{Snapshot}
\vspace{1em}
\begin{equation} \label{eq:snapshot}
    \begin{split}
        [C', M'] := snapshot(tree) {}& \Implies \\
        \forall c \in tree, copy(c) \in {}& C' \\
        \forall m \in tree, copy(m) \in {}& M' \\
        \forall f \in tree, f \in {}& F
    \end{split}
\end{equation}

The non-trivial snapshot operation makes a deep-copy by traversing the full tree or sub-tree of
containers and content that is to be taken a snapshot of. As the snapshot operation is performed
upon a tree of objects it results in two new sets (C' and M') which are then added to C and M
respectively, the referred files in the content remain the same as before the snapshot.

@@

\subsection{Interaction of Multiple Accesses} \label{sec:multiple_access}
As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} in
Section~\ref{sec:conseq} there can be race conditions where the last write wins. This could be
solved by locks or transactions, but as updates are not based on each other, the trade-off for this
inconsistency in favour of scalability and speed is intentional. However, as can also be seen in
consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
the set of files, even ough its modification process is the last one to write to the content.
This is necessary as a file can be referenced from several content and the removal process can not
atomically ensure that the file is not referred to by another content. This results in a lot of
files in the system which are not referenced from anywhere. This can be solved by either making each
file keep track of how many times it is referred to and also make it possible to check this number
and delete the file in an atomic fashion. An easier approach would be to have a garbage collector
which locks parts of the system momentarily meanwhile it removes files which are not referenced.

\par
There is also the chance to have content and containers that are not referenced from anywhere, these
containers and content are defined as garbage. This could happen as a result of
consequence~\ref{eq:delete} in combination with any of the other consequences listed in
Section~\ref{sec:conseq}. The basic idea is that an entity is deleted while another is created or
modified to have that entity as its parent, thus creating a separate tree that can not be reached
from the root container. These entities should be removed, for example by a garbage collector, or
made sure to be deleted in the instance when they are no longer referenced.

\subsection{JPF}
JPF was used to test the idea and state transitions of the model. It is a very simplified
version of the real system that still contains all the important Copy-on-Write concepts and the 
assumptions that have been made in the model. This simplified version was then used to automatically
test for unwanted quirks. It is not a proof that the model works, but it is very exhaustive in its
testing.

@@

\subsubsection{Garbage Collection}
For the system to reach a clean state and eventual consistency~\cite{KLINGSBO}, garbage collection
or reference counters will be needed. The conclusion of Section~\ref{sec:multiple_access} is that
there will be objects in the system that are unreferenced and can not be reached and they should
therefore be cleaned up to not cause negative performance impact on the system.

@@

\subsection{Access Control} \label{sec:disc_access_control}
Full object-level access control was not implemented according to the model described in
Section~\ref{sec:model}. The implementation only checks whether a user should have access to the
system as a whole, and does not set, nor check, specific access rights to contents or containers.

\newpage
\section{Summary}
This section covers the conclusions that I draw from this work and how I plan to extend the
implementation in the future.

@@

\par If this application would be distributed with several persistent storage nodes like MongoDB,
the application would not always be in a global consistent state as there would not be any global
locks. This could theoretically cause some inconvenience for the user where last write
wins~\cite{LASTWRITEWINS} on meta data, but in all real world observations no users have noticed it.
The theoretical inconvenience for the user was a trade-off made so that the application could scale
almost endlessly, especially if the issue with building the project trees is solved (as mentioned in
Section~\ref{sec:load_testing}).

@@

\par The conclusion that can be drawn from implementing Copy-on-Write as high up in the stack as it
was in Perius is that the persistent storage logic would have to be largely rewritten for every
existing application that is to use a similar approach. The solution to this could be to use
Copy-on-Write in the persistent storage instead of in the actual application. This approach could
efficiently handle conflict resolution and every improvement made to it could be shared for every
system that uses the same type of system architecture and underlying persistent storage. On the
other hand, by having all the Copy-on-Write specific code in the back-end instead of the persistent
storage the choices of storage becomes wider as the database does not have to implement
Copy-on-Write and conflict resolution. This could be an advantage when deploying Perius in
environments with high requirements for the number of users that the system should be able to
handle. It would also be well suited if the level of consistency needs to be fully tweakable to make
it possible to trade off consistency for scalability.

\subsection{Future work}
\subsubsection{Access Control}
As access control was not fully implemented (See Section~\ref{sec:disc_access_control}) as planned 
in the model, the implementation will need to be extended in the future.

\subsubsection{Front-end Refactorization}
Currently the implementation fetches the full project tree every time a change is made. This results
in increased latency, bandwidth usage and server load. We propose a refactorization of the front-end
where it updates the state according to the REST response of each modification to the project tree
instead of re-fetching it.

@@
